---
title: "Cyclistic Bike Usage Analysis"
author: "Yt"
date: "2024-08-19"
output:
  html_document: default
  pdf_document: default
---

## Introduction

This document presents an analysis of Cyclistic bike usage, comparing how annual members and casual riders use the bikes differently. The analysis covers usage frequency, ride duration, monthly and daily trends, and provides insights into user behavior.

## Data Preparation

```{r loading packages}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r}
# Verify the working directory
getwd()

# List files in the working directory
list.files()
```
```{r importing data}
tripdata_202212 <- read.csv("/Users/yitingcheng/Documents/Certificates/Google/Basic/8. Case Study/Case Study 1/Raw Data/202212-divvy-tripdata.csv")
tripdata_202301 <- read.csv("/Users/yitingcheng/Documents/Certificates/Google/Basic/8. Case Study/Case Study 1/Raw Data/202301-divvy-tripdata.csv")
tripdata_202302 <- read.csv("/Users/yitingcheng/Documents/Certificates/Google/Basic/8. Case Study/Case Study 1/Raw Data/202302-divvy-tripdata.csv")
tripdata_202303 <- read.csv("/Users/yitingcheng/Documents/Certificates/Google/Basic/8. Case Study/Case Study 1/Raw Data/202303-divvy-tripdata.csv")
tripdata_202304 <- read.csv("/Users/yitingcheng/Documents/Certificates/Google/Basic/8. Case Study/Case Study 1/Raw Data/202304-divvy-tripdata.csv")
tripdata_202305 <- read.csv("/Users/yitingcheng/Documents/Certificates/Google/Basic/8. Case Study/Case Study 1/Raw Data/202305-divvy-tripdata.csv")
tripdata_202306 <- read.csv("/Users/yitingcheng/Documents/Certificates/Google/Basic/8. Case Study/Case Study 1/Raw Data/202306-divvy-tripdata.csv")
tripdata_202307 <- read.csv("/Users/yitingcheng/Documents/Certificates/Google/Basic/8. Case Study/Case Study 1/Raw Data/202307-divvy-tripdata.csv")
tripdata_202308 <- read.csv("/Users/yitingcheng/Documents/Certificates/Google/Basic/8. Case Study/Case Study 1/Raw Data/202308-divvy-tripdata.csv")
tripdata_202309 <- read.csv("/Users/yitingcheng/Documents/Certificates/Google/Basic/8. Case Study/Case Study 1/Raw Data/202309-divvy-tripdata.csv")
tripdata_202310 <- read.csv("/Users/yitingcheng/Documents/Certificates/Google/Basic/8. Case Study/Case Study 1/Raw Data/202310-divvy-tripdata.csv")
tripdata_202311 <- read.csv("/Users/yitingcheng/Documents/Certificates/Google/Basic/8. Case Study/Case Study 1/Raw Data/202311-divvy-tripdata.csv")
```

```{r combine data}
# Combine all dataset
all_trips <- bind_rows(
  tripdata_202212,
  tripdata_202301,
  tripdata_202302,
  tripdata_202303,
  tripdata_202304,
  tripdata_202305,
  tripdata_202306,
  tripdata_202307,
  tripdata_202308,
  tripdata_202309,
  tripdata_202310,
  tripdata_202311
)
```

## Data Processing

```{r data processing}
# Convert data types and create new columns
all_trips <- all_trips %>%
  mutate(
    started_at = ymd_hms(started_at),
    ended_at = ymd_hms(ended_at),
    date = as.Date(started_at),
    year = format(date, "%Y"),
    month = format(date, "%m"),
    day = format(date, "%d"),
    day_of_week = format(date, "%A"),
    ride_length = as.numeric(difftime(ended_at, started_at, units = "mins"))
  ) %>%
  filter(ride_length > 0 & ride_length <= 1440) %>%
  drop_na()
```

## Analysis

### Summary Statistics by User Type
```{r}
summary_by_type <- all_trips %>%
  group_by(member_casual) %>%
  summarise(
    Total_Rides = n(),
    Avg_Ride_Length = mean(ride_length, na.rm = TRUE),
    Median_Ride_Length = median(ride_length, na.rm = TRUE),
    Min_Ride_Length = min(ride_length, na.rm = TRUE),
    Max_Ride_Length = max(ride_length, na.rm = TRUE)
  )

print(summary_by_type)
```

### Total Ride Analysis
#### Monthly Trends
```{r monthly trends}
monthly_trends <- all_trips %>%
  group_by(year, month, member_casual) %>%
  summarise(Total_Rides = n(), .groups = 'drop')

# Line plot
ggplot(monthly_trends, aes(x = as.Date(paste(year, month, "01", sep = "-")), y = Total_Rides, color = member_casual)) +
  geom_line() +
  labs(title = "Monthly Ride Trends by User Type", x = "Month", y = "Total Rides") +
  theme_minimal()

# Bar plot
ggplot(monthly_trends, aes(x = month, y = Total_Rides, fill = member_casual)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Monthly Ride Trends by User Type", x = "Month", y = "Total Rides") +
  theme_minimal()
```

#### Day of Week Trends
```{r day of week trends}
day_of_week_trends <- all_trips %>%
  group_by(day_of_week, member_casual) %>%
  summarise(Total_Rides = n(), .groups = 'drop')

ggplot(day_of_week_trends, aes(x = day_of_week, y = Total_Rides, fill = member_casual)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Total Rides by Day of Week and User Type", x = "Day of Week", y = "Total Rides") +
  theme_minimal()
```

### Ride Length Analysis
#### Monthly Ride Length 
```{r monthly ride length}
monthly_ride_length <- all_trips %>%
  group_by(year, month, member_casual) %>%
  summarise(average_ride_length = mean(ride_length), .groups = 'drop')

# Line plot
ggplot(monthly_ride_length, aes(x = as.Date(paste(year, month, "01", sep = "-")), y = average_ride_length, color = member_casual)) +
  geom_line() +
  labs(title = "Monthly Ride Length by User Type", x = "Month", y = "Average Ride Length") +
  theme_minimal()

# Bar plot
ggplot(monthly_ride_length, aes(x = month, y = average_ride_length, fill = member_casual)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Monthly Ride Length by User Type", x = "Month", y = "Average Ride Length") +
  theme_minimal()
```

#### Daily Ride Length 
```{r daily ride length}
daily_ride_length <- all_trips %>%
  group_by(day_of_week, member_casual) %>%
  summarise(average_ride_length = mean(ride_length), .groups = 'drop')

ggplot(daily_ride_length, aes(x = day_of_week, y = average_ride_length, fill = member_casual)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Average Ride Length by Day of Week and User Type", x = "Day of Week", y = "Average Ride Length") +
  theme_minimal()
```

## Conclusion
1. Usage Frequency
Annual Members: Use Cyclistic bikes significantly more frequently than casual riders, indicating consistent and regular usage.
Casual Riders: Have fewer rides overall, suggesting less frequent use, possibly for specific occasions.
2. Ride Duration
Annual Members: Generally take shorter rides, averaging 12.0 minutes, indicating usage for commuting or shorter trips.
Casual Riders: Tend to have longer rides on average, averaging 20.5 minutes, suggesting leisure or longer trips.
3. Monthly Trends
Annual Members: Show more consistent usage patterns throughout the year.
Casual Riders: Exhibit more fluctuations in ride volume, with noticeable peaks and troughs.
4. Day of the Week Trends
Annual Members: Higher ride volumes on weekdays, especially Tuesday and Thursday.
Casual Riders: Prefer weekends, particularly Saturdays and Sundays.
5. Monthly Average Ride Lengths
Annual Members: Have stable, shorter average ride lengths.
Casual Riders: Show varying, longer average ride lengths.
6. Daily Average Ride Lengths
Annual Members: Maintain consistent average ride lengths throughout the week.
Casual Riders: Experience longer average ride lengths on weekends.

## Summary
Annual members use Cyclistic bikes more frequently and for shorter, more utilitarian trips, often during weekdays. Casual riders use the bikes less frequently but for longer trips, especially on weekends. These differences can inform targeted marketing strategies and operational adjustments.


